---
title: "Assigment 4"
output:
  html_document:
    df_print: paged
date: "2024-10-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
```{r}
# Load necessary libraries
library(readr)
library(dplyr)
library(tidyr)

# Read the CSV file (use `skip = 1` to skip the first empty row)
arrival_delays <- read_csv("/Users/aribarazzaq/Desktop/A_D.csv")

# Check the column names
names(arrival_delays)

# View the first few rows of the dataset to check if it is loaded correctly
head(arrival_delays)

# View the column names
names(arrival_delays)

# Fill in the missing airline names
arrival_delays <- arrival_delays %>%
  fill(...1, .direction = "down") # Replace '...1' if the column name is different

# Pivot to long format: transform destination columns into rows
arrival_delays_long <- arrival_delays %>%
  pivot_longer(cols = starts_with("Los Angeles"), # Adjust based on your actual destination column names
               names_to = "Destination",          # This creates a new 'Destination' column
               values_to = "Flights") %>%         # This contains the flight numbers (delayed or on-time)
  rename(Airline = ...1, Status = ...2) %>%       # Rename the columns for better readability
  filter(!is.na(Flights))                         # Filter out any empty rows

# View the tidy dataset
head(arrival_delays_long)

# Summarize on-time and delayed flights for each airline
summary_delays <- arrival_delays_long %>%
  group_by(Airline, Status) %>%
  summarize(Total_Flights = sum(Flights))

# View the summary
print(summary_delays)




```

```{r}
# Load ggplot2 for plotting
library(ggplot2)

# Create a bar plot comparing on-time and delayed flights by destination and airline
ggplot(arrival_delays_long, aes(x = Destination, y = Flights, fill = Status)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~ Airline) +
  labs(title = "Comparison of On-time and Delayed Flights by Airline",
       x = "Destination",
       y = "Number of Flights")

```
Alaska Airlines had 3274 flights which were  on-time where as 501 were delayed.
This gives Alaska airline a good on-time performance ratio.
AM West Airlines had 6438 flights which were on-time where as 787 were delayed, Eventhough AM West has more number of  flights, the ratio of on-time to delayed flights is not as favorable as Alaska's

So Alaska airline will be my choice since it is on time. 




