---
title: "Lab 02 607"
output: html_document
date: "2024-09-15"
---
#Part 1: Build Table
• Choose six recent popular movies.
#I choose following movies: Barbie, Spider Man, Mission Impossible, Super Mario, Garden of Galaxy

• Ask at least five people that you know (friends, family, classmates, imaginary friends if
necessary) to rate each of these movies that they have seen on a scale of 1 to 5.

#Part 2: Store data in SQL database
 Take the results (observations) and store them in the class MySQL database:
- Server name: cunydata607sql.mysql.database.azure.com
- Username / password: will be given to you in an email
Note: it is good practice to change your password. To do so, use this SQL command:
SET PASSWORD = ‘<your new password here>';


#Part 3: Transfer data from SQL database to R dataframe
 Load the information from the SQL database into an R dataframe.

```{r setup, include=FALSE}
library(DBI)
library(RMySQL)  # Or library(RMariaDB)

# Database connection details
con <- dbConnect(RMySQL::MySQL(), 
                 dbname = "ariba.razzaq96", 
                 host = "cunydata607sql.mysql.database.azure.com", 
                 user = "ariba.razzaq96", 
                 password = "44d86955b2c38238")

# Query the data from the movie_ratings table
query <- "SELECT * FROM `ariba.razzaq96`.`movies rating`;"
movie_data <- dbGetQuery(con, query)

# Display the dataframe
print(movie_data)

# Close the connection when done
dbDisconnect(con)


```
#Part 4: Missing data strategy
 Implement an approach to missing data
 Explain why you decided to take the chosen approach: 
 I chooose to replace null with 0 first because my rating is from 1-5 after that we replaced 0 with mean of the column which is mean imputation. 


```{r}
# Replace missing values (NA) with the column mean
for (col in names(movie_data)[-1]) {
    movie_data[[col]][is.na(movie_data[[col]])] <- mean(movie_data[[col]], na.rm = TRUE)
}

# Check the imputed data
print(movie_data)

```

```{r}

  
  # Calculate the mean of the column, ignoring 0s in the mean calculation
  column_mean <- mean(movie_data[[col]][movie_data[[col]] != 0], na.rm = TRUE)
  
  # Round the mean to the nearest whole number
  rounded_mean <- round(column_mean)
  
  # Replace 0s with the rounded column mean
  movie_data[[col]][movie_data[[col]] == 0] <- rounded_mean


# Check the modified data
  print(movie_data)
```
