---
title: "Project 2 a"
output: html_document
---

Introduction:

https://data.wa.gov/education/Childcare-Need-Supply-All-/hiqz-y2vv/data_preview

The Childcare Need and Supply Dataset provides information on the demand and supply of childcare services across various geographic locations, age groups, and income brackets. The dataset offers key metrics such as:

Subsidized care, private care estimates, and unserved children.
Percent of childcare needs met, segmented by income bracket and age group.


```{r}
# Load the necessary libraries
library(tidyverse)

# Load the dataset
childcare_data <- read.csv("/Users/aribarazzaq/Desktop/Childcare_Need___Supply__All__20241012.csv")

# Preview the structure
str(childcare_data)

```
```{r}
# Rename columns for better readability
childcare_data <- childcare_data %>%
  rename(
    geographic_unit = Geographic.Unit,
    geographic_id = Geographic.ID,
    geographic_name = Geographic.Name,
    income_bracket = State.Median.Income.Bracket,
    age_group = Age.Group,
    subsidized_care = Childcare.Subsidized,
    private_care = Private.Care.Estimate,
    children_receiving_care = Estimated.Children.Receiving.Childcare,
    unserved_children = Estimate.of.Unserved,
    percent_need_met = Percent.Need.Met
  )

```

```{r}
# Check for missing values
sum(is.na(childcare_data))

# Drop rows with missing values, if any
childcare_data <- childcare_data %>% drop_na()

```

```{r}
# Transform care-related metrics into long format
childcare_long <- childcare_data %>%
  pivot_longer(
    cols = subsidized_care:percent_need_met,
    names_to = "care_type",
    values_to = "value"
  )

# Preview the long format data
head(childcare_long)

```

```{r}
# Summarize care metrics by age group
age_group_summary <- childcare_long %>%
  group_by(age_group, care_type) %>%
  summarise(avg_value = mean(value, na.rm = TRUE))

print(age_group_summary)

```

```{r}
# Create a bar plot of average percent need met by income bracket
ggplot(childcare_long %>% filter(care_type == "percent_need_met"), 
       aes(x = income_bracket, y = value, fill = income_bracket)) +
  geom_bar(stat = "summary", fun = "mean") +
  labs(title = "Average Percent of Childcare Need Met by Income Bracket",
       x = "Income Bracket", y = "Percent Need Met") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
```{r}
# Identify age groups with the most unserved children
unserved_summary <- childcare_data %>%
  group_by(age_group) %>%
  summarise(total_unserved = sum(unserved_children, na.rm = TRUE)) %>%
  arrange(desc(total_unserved))

print(unserved_summary)

```

Age Groups: The age groups with the highest number of unserved children can be identified, helping policymakers target resources.
Income Brackets: There are clear disparities in childcare access, with lower-income families having lower percent needs met.
Data Transformation: Pivoting to a long format allowed us to analyze the data more flexibly.

