---
title: "Project 2"
output: html_document
---

Introduction:
This report analyzes the Red Wine Quality dataset to explore how different chemical properties relate to the quality of the wine. Data tidying and transformations were performed using `tidyr` and `dplyr`.
here is the link for the dataset:
https://archive.ics.uci.edu/dataset/186/wine+quality




To tidy and transform the Red Wine Quality dataset following actions have been taken:

1. Renamed Columns: Made column names more descriptive.
2. Handled Missing Values: Checked for and removed missing data.
3. Long Format Transformation: transformed data to a long format for further analysis.
4. Data Transformation: Added new variables and summarized key statistics for better insight into wine quality.




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(stringr)

```

```{r}
# Load the required libraries
library(tidyverse)

# Load the red wine dataset with the correct delimiter
wine_red <- read.csv("/Users/aribarazzaq/Desktop/wine+quality/winequality-red.csv", sep = ";")

# Preview the first few rows to understand the structure
head(wine_red)
str(wine_red)

```

```{r}
# Identify missing values
sum(is.na(wine_red))

# If missing values exist, we can remove them or use imputation
wine_red <- wine_red %>% drop_na()

```


```{r}
# Rename columns to remove spaces and make them more readable
wine_red <- wine_red %>%
  rename(
    fixed_acidity = `fixed.acidity`,
    volatile_acidity = `volatile.acidity`,
    citric_acid = `citric.acid`,
    residual_sugar = `residual.sugar`,
    free_sulfur_dioxide = `free.sulfur.dioxide`,
    total_sulfur_dioxide = `total.sulfur.dioxide`
  )

# Confirm new column names
colnames(wine_red)

```
```{r}
# Convert the data to long format for analysis
wine_red_long <- wine_red %>%
  pivot_longer(
    cols = fixed_acidity:alcohol,
    names_to = "characteristic",
    values_to = "value"
  )

# View the long format dataset
head(wine_red_long)

```

```{r}
# Add a column to classify wine quality
wine_red <- wine_red %>%
  mutate(quality_label = ifelse(quality >= 7, "High Quality", "Low Quality"))

# View the dataset with the new column
head(wine_red)

```
```{r}
# Summarize the average fixed acidity for each quality label
acidity_summary <- wine_red %>%
  group_by(quality_label) %>%
  summarise(mean_fixed_acidity = mean(fixed_acidity, na.rm = TRUE))

print(acidity_summary)

```
```{r}
# Filter wines with high alcohol content (> 10%) and group by quality label
high_alcohol_wines <- wine_red %>%
  filter(alcohol > 10) %>%
  group_by(quality_label) %>%
  summarise(count = n())

print(high_alcohol_wines)

```


To tidy and transform the Red Wine Quality dataset:

Renamed Columns: Made column names more descriptive.
Handled Missing Values: Checked for and removed missing data.
Long Format Transformation: Optionally transformed data to a long format for further analysis.
Data Transformation: Added new variables and summarized key statistics for better insight into wine quality.





